Large-scale harmonic patterns clustering

### 2010-03-01 Paper/Paper.tex
improved

\begin{document}
%
\maketitle
%
\begin{abstract}
We present a method for discovering typical patterns in music.
Similar to the \textit{shingle} idea, but done on a larger scale through
Echo Nest API, we cluster harmonic patterns from 43K songs. Patterns are
made of ordered chroma features, and not of summarization like gaussian
mixtures coefficients or NMF weights. We explain
how to operate on such large datasets, we analyze the obained clusters,
and we discuss there use for retrieval and encoding. All code is made
available.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}

@@

Finding a closest match is a tangible goal within a personal collection
of a few thousands songs. It can be used to discover cover songs or
find the seamless transitions in a playlist. However, the closest match
in millions and millions of music excerpts do not help us organize such
an amount of media. Some tasks are still important, for instance cover
recognition, and improvements in $k$nn methods are always welcome.
But it does not the describe the music space, nor does it necessarily
make good recommendations. A similar argument in the recommendation
domain is that precision becomes more important than recall passed a certain
scale of data for many practical uses.

@@

@@

\section{Previous Work}\label{sec:prevwork}
The idea is closely related to the \textit{Shingles} described by
Slaney and Casey \cite{Casey2006,Casey2007,Casey2008}. 
LSH: \cite{Datar2004}, LSH implementation: \cite{E2LSH}.

@@

Nearest neighbor for music also includes \cite{Cano2004} and
\cite{Holzapfel2009}.

@@

\section{Data}\label{sec:data}

@@

\subsection{Cowbell Dataset}
We have $3,720,091$ non zero bars, a bar usually contains
$4$ beats.

@@

uspop2002 serves as testing set to measure how well a code book learned on
the Cowbell data set can represent new songs.

@@

\subsection{Pattern length}

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|}
\hline
size & {\bf 1 bar (4)} & {\bf 2 bars (16)} & {\bf 4 beats} & {\bf 7 beats} \\ 
\hline \hline
avg. dist & 1 & 2 & 3 & 4 \\ \hline
something & 5 & 6 & 7 & 8 \\ \hline
\end{tabular}
\end{center}
\caption{{Encoding score for different pattern sizes. We can select the
size of the patterns based on bars. In this case, we put the actual size
of the pattern in parentheses. If the pattern size is based on beats, the
size is the number of beats.}}
\label{tab:psize}
\end{table}

>>>>

\begin{document}
%
\maketitle
%
\begin{abstract}
We present a method for discovering typical patterns in music.
Similar to the \textit{shingle} idea, but done on a larger scale through
Echo Nest API, we cluster harmonic patterns from 43K songs. Patterns are
made of ordered chroma features, and not of summarization like gaussian
mixtures coefficients or NMF weights. We explain
how to operate on such large datasets, we analyze the obained clusters,
and we discuss there use for retrieval and encoding. All code is made
available.\\
TAKE-HOME MESSAGE: \textit{Shingles} idea can go large-scale?
\end{abstract}
%
\section{Introduction}\label{sec:introduction}

@@

Finding a closest match is a tangible goal within a personal collection
of a few thousands songs. It can be used to discover cover songs or
find the seamless transitions in a playlist. However, the closest match
in millions and millions of music excerpts do not help us organize such
an amount of media. Some tasks are still important, for instance cover
recognition, and improvements in $k$nn methods are always welcome.
But it does not the describe the music space, nor does it necessarily
make good recommendations. We do not want to look for one neighbor, but
rather a cloud of potential close patterns. A similar argument in the 
recommendation domain is that precision becomes more important than recall 
passed a certain scale of data for many practical uses.

@@

Our contribution is... large-scale clustering method,

@@

\section{Previous Work}\label{sec:prevwork}
The idea is closely related to the \textit{Shingles} described by
Slaney and Casey \cite{Casey2006,Casey2007,Casey2008}.

LSH: \cite{Datar2004}, LSH implementation: \cite{E2LSH}.
The E$^2$LSH packages can efficiently handle about $50$K or $100$K
vectors of size $200$. Even with a better implementation, the LSH
model still has some scaling problem. A large portion of the data
must be known to decide how many projections to use, and what guarantee
of finding the nearest neighbor it can provides. Also, we could fix the
number of projections and consider as clusters all the elements that
fall into the same bin. However, this does not give us a centroid, a
typical pattern representing that cluster.

@@

Nearest neighbor and clustering methods are common in music similarity 
research, for instance \cite{Cano2004,Holzapfel2009}. The difference here 
are the actual patterns that we cluster.

@@

\section{Data}\label{sec:data}
In this section we present our use of the Echo Nest features and the
datasets we use for training and testing our model.

@@

\subsection{Cowbell Dataset}
We have $43,300$ songs, giving  $3,720,091$ non zero bars, a bar usually 
contains $4$ beats.

@@

uspop2002 serves as testing set to measure how well a code book learned on
the Cowbell data set can represent new songs. We get Echo Nest features
for $8651$ songs from that dataset.

@@

\subsection{Pattern length}
See encoding experiments in Table \ref{tab:psize}.

% RESULTS
% 200codes, 2 bars, pSize=16, maxener, True, False, not scaled = 3.788
%                                             averaged on size = 
% 200codes, 2 bars, pSize=8, maxener, True, False, not scaled  = 3.135
%                                             averaged on size = 
% 200codes, 1 bars, pSize=4, maxener, True, False, not scaled  = 
%                                             averaged on size = 
% 200codes, 7 beats, pSize=7, maxener, True, False, not scaled = 2.771
%                                             averaged on size = 
% 200codes, 5 beats, pSize=5, maxener, True, False, not scaled = 2.397
%                                             averaged on size = 
% 200codes, 3 beats, pSize=3, maxener, True, False, not scaled = 
%                                             averaged on size = 
\begin{table}
\begin{center}
\begin{tabular}{|l|c|}
\hline
size & avg. dist \\ \hline \hline
1 bar (4) & \\
2 bars (8) & $$\\
2 bars (12) & $$\\
2 bars (16) & $$ \\ \hline
3 beats & \\
5 beats & $$\\
7 beats & $$\\ \hline
\end{tabular}
\end{center}
\caption{{Encoding score for different pattern sizes. We can select the
size of the patterns based on bars. In this case, we put the actual size
of the pattern in parentheses. If the pattern size is based on beats, the
size is the number of beats. Results are on the uspop dataset, not seen
during training.}}
\label{tab:psize}
\end{table}

\subsection{Data Size}
We are interesting to see how the amount of training data changes
the quality of our code.
We experiment with different data sizes for a fixed size of codebook and a 
fixed code length. The data points are chose randomly from the whole
data set.
Code length: 1 bar resized to size 4. Code book size: 100
We measure the distortion on our test set. We hope for some convergence.
See results in Table \ref{tab:datasize}.

Our learning can overfit, which explain the bad scores of small datasets.
The larger the dataset, the smallest the chance to overfit.
We ar elooking for a size of dataset that guarantees us a good learning
process, while being unnecessary large.

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
$10$K & $100$K & $500$K & $2$M  \\ \hline
$0$ & $1$ & $2$ & $3$\\ \hline
\end{tabular}
\end{center}
\caption{{Distortion for a code book of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{tab:datasize}
\end{table}
% RESULTS:    avg.       mean dist
% 10K         
% 100K
% 500K
% 2M

\subsection{Codebook Size in Function of Pattern Length}
If the signal was totally random, and we were using 2 code to encode
every beat, we would need.... 4 codes for 2 beats? and 16 codes for
4 beats? and 256 codes for 8 beats?

Think about it in binary: two code to encode (0) and (1), but for
(0,0),(0,1),(1,0),(1,1) we need four codes for the same accuracy.

Reasonable size: 2 - 4 - 16 - 256. If we improve distance, music
is not random!

Then... 3 beats, 5 beats, 7 beats don't encode as well!

Special: 4 beats versus 1 bar - 4 beats, show the problem of being
not in sync with bars.
